\chapter{Graphs}
\section{Graphs}
\begin{definition}
  A \emph{graph} is a pair
  \begin{equation*}
    G = (V, E),
  \end{equation*}
  where $V$ is a finite set of \emph{vertices}, and $E$ is a set of
  \emph{edges}, which are 2-element subsets of $V$.
  \begin{itemize}
    \item The elements of an edge is called its \emph{endpoints}, and we say
    that an edge \emph{joins} its endpoints.
    \item We say that two vertices are \emph{adjacent} if they are joined by an
    edge.
  \end{itemize}
\end{definition}
\begin{remark}
  An edge $\{u, v\}$ is usually written as $uv$.
\end{remark}

\section{Paths}
\begin{definition}
  A \emph{walk} of a graph $G = (V, E)$ is a sequence
  \begin{equation*}
    (v_0, e_1, v_1, \dots, e_k, v_k),
  \end{equation*}
  where $v_0, v_1, \dots, v_k \in V$ and $e_1, \dots, e_k \in E$ such that
  $e_i = v_{i-1}v_i$ for each $i \in \{1, \dots, k\}$.
  The number of edges in a walk is called the \emph{length} of the walk.
  \begin{itemize}
    \item A \emph{trail} is a walk with no repeated edges.
    \item A \emph{circuit} is a trail with starting vertex and ending vertex
    being the same.
    \item A \emph{path} is a walk with no repeated vertices.
    \item A \emph{cycle} is a circuit with no repeated internal vertices.
  \end{itemize}
\end{definition}

\section{Connectivity}
\begin{definition}
  A graph is \emph{connected} if any two vertices can be linked by a path.
  A graph is \emph{disconnected} if it is not connected.
\end{definition}

\begin{definition}
  Let $G = (V, E)$ be a graph.
  We say that a set $S \subseteq V$ \emph{separates} $G$ if $G[V \setminus S]$
  is disconnected.
\end{definition}

\begin{definition}
  Let $G = (V, E)$ be a graph and let $k$ be a nonnegative integer.
  We say that $G$ is \emph{$k$-connected} if $|V| > k$ and every subset
  $S$ of $V$ with $|S| < k$ does not separate $G$.
\end{definition}